SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-11-30 abs.tex
made abstract more "punchy" and direct

\begin{abstract}
Hardware prefetching has been introduced in modern processors as a way to hide
cache latencies.  An efficient prefetcher should be able to identify complex
memory access patterns during program execution. This ability enables the
prefetcher to read a block ahead of its demand access, potentially saving a
cache miss. Accurately identifying the right blocks to prefetch is essential
to achieving high performance from the prefetcher.

In this paper, we introduce Perceptron-based Prefetch Filtering to help make
this prefetching decision accurately. The perceptron layer acts as a check 
to filter out the unnecessary prefetches recommended by the underlying
prefetcher.  We have also explored a range of features that can be used to
train the perceptron layer.  Our results show that perceptron-based filtering
improves performance on the memory intensive subset of the SPEC 2017 benchmark
suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared
to a state-of-the art prefetcher.  We also demonstrate that the performance
gained from using our efficient filter scales better with increasing numbers
of cores.

>>>>

\begin{abstract}

Hardware prefetching is an effective technique for hiding cache miss latencies
in modern processor designs.  Prefetcher performance can be characterized by 
two main metrics, which are generally at odds with one another: coverage 
(the improvement in the cache miss metric) and accuracy (the fraction of 
useful prefetches).  An overly aggressive prefetcher may improve coverage, but
may waste so many resources (e.g., bandwidth) that performance is ultimately 
harmed.  An ideal prefetcher would have both very high coverage and accuracy.

%Hardware prefetching has been introduced in modern processors as a way to hide
%cache latencies.  An efficient prefetcher should identify complex
%memory access patterns during program execution. This ability enables the
%prefetcher to read a block ahead of its demand access, potentially preventing a
%cache miss. Accurately identifying the right blocks to prefetch is essential
%to achieving high performance from the prefetcher.

In this paper, we introduce Perceptron-based Prefetch Filtering (PPF) as a way
to reject inaccurate candidate prefetches generated by a baseline prefetcher.  
PPF directly increases prefetcher accuracy, and enables more aggressive tuning
of the baseline prefetcher, which in turn leads to increased coverage.  We also
explore a range of features to use to train PPF's perceptron layer.  Our results
show that PPF improves performance on a memory-intensive subset of SPEC CPU 2017
benchmarks by 6.84\% for a single-core configuration, and by 11.9\% for a 4-core
configuration, compared to the baseline prefetcher alone.  We also show that
performance continues to scale as the number of cores sharing a last level cache
increases.

%In this paper, we introduce Perceptron-based Prefetch Filtering to help make
% prefetching decisions accurately. The perceptron layer acts as a check 
%to filter out the unnecessary prefetches recommended by the underlying
%prefetcher.  We have also explored a range of features that can be used to
%train the perceptron layer.  Our results show that perceptron-based filtering
%improves performance on the memory intensive subset of the SPEC 2017 benchmark
%suite by 6.84\% on single-core and by 11.9\% on multi-core traces, as compared
%to a state-of-the art prefetcher.  We also demonstrate that the performance
%gained from using our efficient filter continues to scale as the number of cores sharing a 
%last level cache increases.
